---
title: "00-Weather"
author: "Danielle Ethier"
date: "06/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}

require(RNCEP) #for accessing the weather data
require(tidyverse) 
require(lubridate)

```


## Weather data access

Data come from the NCEP/NCAR Reanalysis (http://www.esrl.noaa.gov/psd/data/gridded/data.ncep.reanalysis.html) 
and NCEP/DOE Reanalysis II (http://www.esrl.noaa.gov/psd/data/gridded/data.ncep.reanalysis2.html) datasets. 

Note that variables on a T62 Gaussian grid are evenly spaced in longitude but unevenly spaced in latitude. All data are downloaded on the gaussian grid for compariative purposes. 
  
Date range 1995-2020
Month range July-October (7,10)
Lat Long coordinates cover the Great Lakes Shoreline associated with MOBU counts from Long Point and surrounding area

Zipf 2017: Total Monthly Precipitation, Average Monthly Temperature. 

Bounding box: northeast Ottawa to tip of the Bruce in the northwest, down to LPBO in the south 

#Air temp monthly average RNCEP
```{r temp}

#Import saved summary data
#air_temp_summary<-read.csv("Mean.Monthly.Temp.csv")

air_temp <- NCEP.gather(variable = 'air.2m', level = 'gaussian', months.minmax = c(6, 10), years.minmax = c(1995,2020), lat.southnorth = c(42.50, 45.40), lon.westeast = c(-81.65, -75.69), return.units = TRUE)

#Calculate the mean air temperature on a given day. 
air_temp_mean <- NCEP.aggregate(wx.data=air_temp, HOURS=FALSE, fxn='mean')

#Change data from an array into a dateframe
air_temp_mean <- NCEP.array2df(air_temp_mean, var.names=NULL)

#Change temperature into degree C and create year, month, day columns
air_temp_summary <- air_temp_mean %>%
         mutate(air_temp_mean = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1) %>%
         group_by(month_year) %>%
         dplyr::summarize(air_temp_mean=mean(air_temp_mean)) %>%
	       separate(month_year, into=c("Year", "month"), sep=" ") %>%
         pivot_wider(names_from = month, values_from = air_temp_mean) %>%
         rename(June_temp = "6", July_temp = "7", August_temp = "8", September_temp = "9", October_temp = "10") %>%
         mutate(Year=as.numeric(Year)) %>% 
         mutate (Summer_temp = rowMeans(select(., 3:5))) #average summer temp June-Aug

air_temp_summary_vis <- air_temp_mean %>%
         mutate(air_temp_mean = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1) %>%
         group_by(month_year) %>%
         dplyr::summarize(air_temp_mean=mean(air_temp_mean)) %>%
	 separate(month_year, into=c("year", "month"), sep=" ") %>%
         mutate(year = as.numeric(year), month = as.numeric(month))

ggplot(air_temp_summary_vis, aes(x=year, y=air_temp_mean)) +
  geom_point() +
  geom_smooth() +
  theme_classic() +
  facet_wrap(month ~., scales="free")

#save temperature data
write.csv(air_temp_summary,"Mean.Monthly.Temp.csv")
air_temp_summary<-read.csv("Mean.Monthly.Temp.csv")

```  


#Total (sum) monthly precipitation RNCEP
```{r precipitation}

#Import saved summary data
#precip_summary<-read.csv("Sum.Monthly.Precip.csv")

precip <- NCEP.gather(variable = 'prate.sfc', level = 'gaussian', months.minmax = c(6, 10), years.minmax = c(1995,2020), lat.southnorth = c(42.50, 45.40), lon.westeast = c(-81.65, -75.69), return.units = TRUE)

precip_sum<- NCEP.aggregate(wx.data=precip, HOURS=FALSE, fxn='sum')

precip_sum <- NCEP.array2df(precip_sum, var.names=NULL)

precip_summary <- precip_sum %>%
         mutate(precip_mean = variable1,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1) %>%
         group_by(month_year) %>%
         dplyr::summarize(precip_mean=mean(precip_mean)) %>%
	 separate(month_year, into=c("Year", "month"), sep=" ") %>%
         pivot_wider(names_from = month, values_from = precip_mean) %>%
         rename(June_precip = "6", July_precip = "7", August_precip = "8", September_precip = "9", October_precip = "10") %>%
         mutate(Year = as.numeric(Year)) %>% 
         mutate (Summer_precip = rowMeans(select(., 3:5))) #average summer precip June-Aug

precip_summary_vis <- precip_sum %>%
         mutate(precip_mean = variable1,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1) %>%
         group_by(month_year) %>%
         dplyr::summarize(precip_mean=mean(precip_mean)) %>%
	 separate(month_year, into=c("year", "month"), sep=" ") %>%
         mutate(year = as.numeric(year), month = as.numeric(month))


ggplot(precip_summary_vis, aes(x=year, y=precip_mean)) +
  geom_point() +
  geom_smooth() +
  theme_classic() +
  facet_wrap(month ~., scales="free")


#save precipitation data
write.csv(precip_summary,"Sum.Monthly.Precip.csv")
precip_summary <-read.csv("Sum.Monthly.Precip.csv")
```


##Combine weather covariates into one table

```{r combine weather}

cov<-merge(air_temp_summary, precip_summary, by=c("Year"))

```
