---
title: "00-Weather v2"
author: "Danielle Ethier"
date: "26/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install required packages

```{r installPackages, message = FALSE, warning = FALSE}

require(tidyverse)
require(RNCEP) #for accessing the weather data
require(climwin)
require(lubridate)
require(pollen) #calculate growing degree days (gdd)

```

## Weather data access

Data come from the NCEP/NCAR Reanalysis (http://www.esrl.noaa.gov/psd/data/gridded/data.ncep.reanalysis.html) 
and NCEP/DOE Reanalysis II (http://www.esrl.noaa.gov/psd/data/gridded/data.ncep.reanalysis2.html) datasets. 

Note that variables on a T62 Gaussian grid are evenly spaced in longitude but unevenly spaced in latitude. All data are downloaded on the gaussian grid for compariative purposes. 
  
Date range 1995-2020
Month range July-October (7,10)
Lat Long coordinates cover the Great Lakes Shoreline associated with MOBU counts from Long Point and surrounding area

Zipf 2017: Total Monthly Precipitation, Average Monthly Temperature, Average min daily temp 

Bounding box: northeast Ottawa to tip of the Bruce in the northwest, down to LPBO in the south 

#Air temp daily mean
```{r tempmin}

air <- NCEP.gather(variable = 'tmin.2m', level = 'gaussian', months.minmax = c(6, 10), years.minmax = c(1995,2020), lat.southnorth = c(42.50, 45.40), lon.westeast = c(-81.65, -75.69), return.units = TRUE)

#Calculate the mean min air temperature on a given day. 
air_mean <- NCEP.aggregate(wx.data=air, HOURS=FALSE, fxn='mean')

#Change data from an array into a dateframe
air_mean <- NCEP.array2df(air_mean, var.names=NULL)

air_mean <- air_mean %>%
         mutate(air_mean = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime))) %>% 
         group_by(datetime) %>%
         dplyr::summarize(air_mean=mean(air_mean))

air_mean$date <- format(air_mean$datetime, format="%d/%m/%Y")
air_mean$date<-as.factor(air_mean$date)

write.csv(air_mean, "Daily mean air temp.csv")
```  

#Air temp daily min
```{r tempmin}

#Calculate the mean min air temperature on a given day. 
air_min <- NCEP.aggregate(wx.data=air, HOURS=FALSE, fxn='min')

#Change data from an array into a dateframe
air_min <- NCEP.array2df(air_min, var.names=NULL)

air_min_mean <- air_min %>%
         mutate(min_temp = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)))%>% 
         select(-variable1) %>% 
         group_by(datetime) %>%
         dplyr::summarize(air_min=mean(min_temp))

air_min_mean$date <- format(air_min_mean$datetime, format="%d/%m/%Y")
air_min_mean$date<-as.factor(air_min_mean$date)

write.csv(air_min_mean, "Daily min air temp.csv")
``` 

#Air temp daily max
```{r tempmin}

#Calculate the mean min air temperature on a given day. 
air_max <- NCEP.aggregate(wx.data=air, HOURS=FALSE, fxn='max')

#Change data from an array into a dateframe
air_max <- NCEP.array2df(air_max, var.names=NULL)

air_max_mean <- air_max %>%
         mutate(max_temp = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)))%>% 
         select(-variable1) %>% 
         group_by(datetime) %>%
         dplyr::summarize(air_max=mean(max_temp))

air_max_mean$date <- format(air_max_mean$datetime, format="%d/%m/%Y")
air_max_mean$date<-as.factor(air_max_mean$date)

write.csv(air_max_mean, "Daily max air temp.csv")
``` 

#Total (sum) daily precipitation RNCEP
```{r precipitation}

precip <- NCEP.gather(variable = 'prate.sfc', level = 'gaussian', months.minmax = c(6, 10), years.minmax = c(1995,2020), lat.southnorth = c(42.50, 45.40), lon.westeast = c(-81.65, -75.69), return.units = TRUE)

precip_sum<- NCEP.aggregate(wx.data=precip, HOURS=FALSE, fxn='sum')

precip_sum <- NCEP.array2df(precip_sum, var.names=NULL)

precip_sum <- precip_sum %>%
         mutate(precip_tot = variable1,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime))) %>%
         select( -variable1) %>% 
         group_by(datetime) %>%
         dplyr::summarize(precip_sum=mean(precip_tot))

precip_sum$date <- format(precip_sum$datetime, format="%d/%m/%Y")
precip_sum$date<-as.factor(precip_sum$date)

#save precipitation data
write.csv(precip_sum,"Daily tot precip.csv")

```

```{r merge weather}

cov<-merge(air_mean, air_max_mean, by=c("datetime", "date"))
cov<-merge(cov, air_min_mean, by=c("datetime", "date"))
cov<-merge(cov, precip_sum, by=c("datetime", "date"))

```

#Caculate growing degree days and daily temperature varaibility (difference between daily high and low)

```{r GDD}
cov<-cov %>% mutate(doy=yday(datetime))

cov$gdd<-gdd(tmax=cov$air_max, tmin=cov$air_min, tbase=11, tbase_max = 33, type="D")

cov<-cov %>% mutate(air_var=air_max-air_min)

```


```{r write weather}

write.csv(cov, "Daily weather.csv")

```

