---
title: "03-SupplementalAnalysis"
author: "Danielle Ethier"
date: "09/07/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install required packages

```{r installPackages}

require(tidyverse)  
require(psych)
require(mgcv)
require(lmtest)
require(RNCEP)

detach("package:plyr", unload=TRUE) #if this was previously loaded it will cause a headache. 
 
```

##Import data and select desired covariates

```{r import data}

dat<-read.csv("Monarch.data.2021.csv")

dat<-dat %>% select(survey_year, June_temp, July_temp, August_temp, September_temp, October_temp, Summer_temp, Summer_temp, June_precip, July_precip, August_precip, September_precip, October_precip, Summer_precip)

```

#Model Weather with Time (year)

Look to see if there has been a significant linear change in weather over time. Hint: there is no diserable pattern in any of our weather covariates. 

```{r weather model}

JuneT<-lm(June_temp~survey_year, data=dat)
summary(JuneT)

JulyT<-lm(July_temp~survey_year, data=dat)
summary(JulyT)

AugT<-lm(August_temp~survey_year, data=dat)
summary(AugT)

SeptT<-lm(September_temp~survey_year, data=dat)
summary(SeptT)

OctT<-lm(October_temp~survey_year, data=dat)
summary(OctT)

JuneP<-lm(June_precip~survey_year, data=dat)
summary(JuneP)

JulyP<-lm(July_precip~survey_year, data=dat)
summary(JulyP)

AugP<-lm(August_precip~survey_year, data=dat)
summary(AugP)

ggplot(dat, aes (x=survey_year, y=August_precip)) +
  geom_point() +
  geom_smooth(method ="lm") + 
  xlab("Survey Year") +
  ylab("August Precipitation") +
  theme_classic()

SeptP<-lm(September_precip~survey_year, data=dat)
summary(SeptP)

OctP<-lm(October_precip~survey_year, data=dat)
summary(OctP)

```


#Spatial autocorrelation of environmental covariates

Look to see if there is significant spatail autocorrelation in our weather covarites over the spatail scale of the analysis. 

```{r spatial autocorrelation}

air_temp <- NCEP.gather(variable = 'air.2m', level = 'gaussian', months.minmax = c(6, 10), years.minmax = c(1995,2020), lat.southnorth = c(42.50, 45.40), lon.westeast = c(-81.65, -75.69), return.units = TRUE)

#Calculate the mean air temperature on a given day. 
air_temp_mean <- NCEP.aggregate(wx.data=air_temp, HOURS=FALSE, fxn='mean')

#Change data from an array into a dateframe
air_temp_mean <- NCEP.array2df(air_temp_mean, var.names=NULL)

#Change temperature into degree C and create year, month, day columns
air_temp_summary <- air_temp_mean %>%
         mutate(air_temp_mean = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1)%>%
         group_by(month_year, latitude, longitude) %>%
         dplyr::summarize(air_temp_mean=mean(air_temp_mean)) %>%
	       separate(month_year, into=c("Year", "month"), sep=" ") %>%
         pivot_wider(names_from = month, values_from = air_temp_mean) %>%
         rename(June_temp = "6", July_temp = "7", August_temp = "8", September_temp = "9", October_temp = "10") %>%
         mutate(Year=as.numeric(Year))

air_temp_summary_vis <- air_temp_mean %>%
         mutate(air_temp_mean = variable1 - 273,
         longitude = longitude - 360,  
         datetime = ymd(gsub("[_XX]", "", datetime)),
         year = year(datetime),
         month = month(datetime),
         day = mday(datetime),
         month_year = paste(year,month)) %>%
         select(-datetime, -variable1) %>%
         group_by(month_year, latitude, longitude) %>%
         dplyr::summarize(air_temp_mean=mean(air_temp_mean)) %>%
	 separate(month_year, into=c("year", "month"), sep=" ") %>%
         mutate(year = as.numeric(year), month = as.numeric(month))

ggplot(air_temp_summary_vis, aes(x=year, y=air_temp_mean)) +
  geom_point() +
  geom_smooth() +
  theme_classic() +
  facet_wrap(month ~., scales="free")




```
